---
title: "Healthcare Facilities and Risk Factors"
author: "Nathan Seto"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
```{r setup, include = F}
# rm(list = ls())
library(knitr)
library(formatR)
library(tidyverse)
library(readxl)
library(readr)
library(ggcorrplot)
library(ggplot2)
library(leaflet)
library(leaflet.extras)
library(magrittr)
library(sf)
opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
```
```{r functions, include = F}
state_facility_load <- function() {
  # health risks
  health <- read_csv("~/internship/workspace/PLACES__Local_Data_for_Better_Health__County_Data_2024_release_20250306.csv")
  colnames(health) <- tolower(names(health))
  
  states <- c("Colorado", "Montana", "North Dakota", "South Dakota", "Utah", "Wyoming")
  
  health <- filter(health, state %in% states) |> arrange(stateabbr, county) |> select(-locationid)
  
  # health facilities
  facility_files <- list(
    colorado = "~/internship/workspace/Health Facility Data/ColoradoHealth24_xlsx.xlsx",
    wyoming = "~/internship/workspace/Health Facility Data/WyomingHealth24_xlsx.xlsx",
    south_dakota = "~/internship/workspace/Health Facility Data/SouthDakotaHealth24_xlsx.xlsx",
    north_dakota = "~/internship/workspace/Health Facility Data/NorthDakotaHealth24_xlsx.xlsx",
    utah = "~/internship/workspace/Health Facility Data/UtahHealth24_xlsx.xlsx",
    montana = "~/internship/workspace/Health Facility Data/MontanaHealth24_xlsx.xlsx"
  )
  
  facilities <- lapply(names(facility_files), function(state) {
    file <- facility_files[[state]]
    data <- read_excel(file, sheet = ifelse(state %in% c("north_dakota", "montana"), 1, 2))
    
    if (state == "wyoming") {
      data$county <- gsub(" County", "", data$county)
    }
    
    if (state == "montana") {
      data$county <- str_to_title(data$county)
    }
    mhc_count$state <- tolower(mhc_count$state)
    data <- data |> left_join(mhc_count, by = c("state", "county"))
    data$MigrantHealthCenters[is.na(data$MigrantHealthCenters)] <- 0
    
    assign(state, data, envir = .GlobalEnv)
  })
  
  # H2-A population
  h2a_population <- read_csv("~/internship/workspace/Written Datasets/h2a_by_county_new.csv") |> arrange(state, county)
  
  assign("health", health, envir = .GlobalEnv)
  assign("h2a_population", h2a_population, envir = .GlobalEnv)
  assign("states", states, envir = .GlobalEnv)
}


migrant_health_centers <- function() {
  # migrant health centers specifically
  ncfh <- read_csv("~/internship/workspace/migrant_health_centers_ncfh.csv")
  ncfh$county <- str_to_title(ncfh$county)
  ncfh$state <- str_to_title(ncfh$state)
  assign("ncfh", ncfh, envir = .GlobalEnv)
  
  mhc_count <- ncfh |>
    group_by(state, county) |>
    summarise(MigrantHealthCenters = n(), .groups = "drop")
  assign("mhc_count", mhc_count, envir = .GlobalEnv)
}


healthcare <- function() {
  # Helper function to process state health data
  process_state_health <- function(state_name, state_data) {
    state_health <- subset(health, state == state_name)
    state_health <- state_health |>
      left_join(state_data, by = "county") |>
      select(-state.y) |>
      rename(state = state.x) |>
      left_join(h2a_population, by = c("county", "state"))
    
    col_order <- c(head(names(state_health), -2), tail(names(state_health), 2)[2], tail(names(state_health), 2)[1])
    state_health <- state_health[, col_order]
    
    state_health
  }
  
  # Process each state's health data
  co_health <- process_state_health("Colorado", colorado)
  wy_health <- process_state_health("Wyoming", wyoming)
  sdkt_health <- process_state_health("South Dakota", south_dakota)
  ndkt_health <- process_state_health("North Dakota", north_dakota)
  ut_health <- process_state_health("Utah", utah)
  mt_health <- process_state_health("Montana", montana)
  
  # Add to global environment
  assign("co_health", co_health, envir = .GlobalEnv)
  assign("ut_health", ut_health, envir = .GlobalEnv)
  assign("wy_health", wy_health, envir = .GlobalEnv)
  assign("mt_health", mt_health, envir = .GlobalEnv)
  assign("ndkt_health", ndkt_health, envir = .GlobalEnv)
  assign("sdkt_health", sdkt_health, envir = .GlobalEnv)
}


stateFacilityCorrelations <- function() {
  # Initialize all 6 data frames for storing correlation coefficients
  init_corr_df <- function() {
    data.frame(
      Predictor = character(),
      Correlation = numeric(),
      PValue = numeric(),
      Significance = numeric()
    )
  }
  
  co_corr <- init_corr_df()
  mt_corr <- init_corr_df()
  ndkt_corr <- init_corr_df()
  sdkt_corr <- init_corr_df()
  wy_corr <- init_corr_df()
  ut_corr <- init_corr_df()
  
  # Helper function to calculate correlations
  calculate_correlations <- function(health_data, state_name) {
    predictors <- names(health_data[, (which(names(health_data) == "geolocation") + 1):(which(names(health_data) == "total_workers_h2a") - 1)])
    corr_df <- init_corr_df()
    for (predictor in predictors) {
      cor_test <- cor.test(health_data[[predictor]], health_data$total_workers_h2a, method = "spearman", exact = FALSE, use = "pairwise.complete.obs")
      corr_df <- rbind(corr_df, data.frame(
        Predictor = predictor,
        Correlation = round(cor_test$estimate, 3),
        PValue = cor_test$p.value
      ))
    }
    corr_df <- corr_df |>
      mutate(Variable = "total_workers_h2a", .before = Predictor) |>
      mutate(State = state_name, .before = Variable)
    corr_df$Significance <- ifelse(corr_df$PValue <= 0.05, TRUE, FALSE)
    corr_df
  }
  
  # Calculate correlations for each state
  co_corr <- calculate_correlations(co_health, "Colorado")
  mt_corr <- calculate_correlations(mt_health, "Montana")
  ndkt_corr <- calculate_correlations(ndkt_health, "North Dakota")
  sdkt_corr <- calculate_correlations(sdkt_health, "South Dakota")
  wy_corr <- calculate_correlations(wy_health, "Wyoming")
  ut_corr <- calculate_correlations(ut_health, "Utah")
  
  # Binding the data together
  h2aPopulationCorrelations <- rbind(co_corr, wy_corr, ndkt_corr, sdkt_corr, ut_corr, mt_corr) |> dplyr::arrange(State, Correlation)
  assign("h2aPopulationCorrelations", h2aPopulationCorrelations, envir = globalenv())
}


weather_facility_load <- function() {
  # Load and process disaster data
  disaster <- read_csv("~/internship/workspace/Written Datasets/Disaster_clean.csv")
  disaster <- disaster |> select(-grep("coastal|Tsunami|hurricane|volcanic", colnames(disaster), ignore.case = TRUE))
  disaster[is.na(disaster)] <- 0
  colnames(disaster) <- gsub(" ", "", colnames(disaster))
  
  assign("disaster", disaster, envir = globalenv())
  
  # Prepare a list of data frames for each state
  data_frames <- list(colorado, montana, north_dakota, south_dakota, utah, wyoming)
  full_data_list <- list()
  
  # Process data for each state
  for (i in seq_along(states)) {
    state_disaster <- subset(disaster, state == states[i])
    full_data <- left_join(state_disaster, data_frames[[i]], by = 'county') |>
      select(-state.y) |>
      rename(state = state.x) |>
      select(where(is.numeric))
    full_data[is.na(full_data)] <- 0
    full_data_list[[states[i]]] <- full_data
  }
  
  # Add ID column and filter out non-numeric columns
  add_id_and_filter_numeric <- function(df) {
    df |> mutate(id = 1:nrow(df), .before = `Population2020`) |> select(where(is.numeric))
  }
  
  full_data_list <- lapply(full_data_list, add_id_and_filter_numeric)
  
  
  # Assign individual state data frames
  assign("utah_full", full_data_list[["Utah"]], envir = .GlobalEnv)
  assign("colorado_full", full_data_list[["Colorado"]], envir = .GlobalEnv)
  assign("wyoming_full", full_data_list[["Wyoming"]], envir = .GlobalEnv)
  assign("montana_full", full_data_list[["Montana"]], envir = .GlobalEnv)
  assign("northDakota_full", full_data_list[["North Dakota"]], envir = .GlobalEnv)
  assign("southDakota_full", full_data_list[["South Dakota"]], envir = .GlobalEnv)
  
  # Clean column names in the final data frames
  clean_column_names <- function(df) {
    colnames(df) <- gsub("_", "", colnames(df))
    colnames(df) <- gsub("-", "", colnames(df))
    colnames(df) <- gsub(" ", "", colnames(df))
    df
  }
  
  # colorado_full <- colorado_full |> rename(MigrantHealthCenters = migranthealthcenters)
  # utah_full <- utah_full |> rename(MigrantHealthCenters = migranthealthcenters)
  # wyoming_full <- wyoming_full |> rename(MigrantHealthCenters = migranthealthcenters)
  # northDakota_full <- northDakota_full |> rename(MigrantHealthCenters = migranthealthcenters)
  # southDakota_full <- southDakota_full |> rename(MigrantHealthCenters = migranthealthcenters)
  # montana_full <- montana_full |> rename(MigrantHealthCenters = migranthealthcenters)
  
  # Apply cleaning function to each state data frame
  assign("utah_full", clean_column_names(utah_full), envir = .GlobalEnv)
  assign("colorado_full", clean_column_names(colorado_full), envir = .GlobalEnv)
  assign("wyoming_full", clean_column_names(wyoming_full), envir = .GlobalEnv)
  assign("montana_full", clean_column_names(montana_full), envir = .GlobalEnv)
  assign("northDakota_full", clean_column_names(northDakota_full), envir = .GlobalEnv)
  assign("southDakota_full", clean_column_names(southDakota_full), envir = .GlobalEnv)
}


create_response_predictors <- function() {
  # data_frames2 <- list(colorado_full, montana_full, northDakota_full, southDakota_full, utah_full, wyoming_full)
  response_variables <- c("Hospitals",
                          "CommunityClinics",
                          "FreeStandingEmergencyDepartments",
                          "RehabilitationHospitals",
                          "RuralClinics",
                          "RuralHealthClinics",
                          "CriticalAccessHospitals",
                          "Mammography",
                          "HomeHealthAgency",
                          "AssistedLivingFacilityTypeI",
                          "AssistedLivingFacilityTypeII",
                          "EndStageRenalDiseaseFacility",
                          "BirthingCenter",
                          "AbortionClinic",
                          "NursingCareFacility",
                          "SmallHealthCareFacility",
                          "SmallHealthCareFacilityTypeN",
                          "PersonalCareAgency",
                          "MigrantHealthCenters")
  
  predictor_variables <- setdiff(colnames(colorado_full), response_variables)[-c(1:5)]
  
  assign("response", response_variables, envir = .GlobalEnv)
  assign("predictors", predictor_variables, envir = .GlobalEnv)
  
}


healthFacility_envir_correlations <- function(data, response, predictors) {
  corr_df <- data.frame(
    Predictor = character(),
    Response = character(),
    Correlation = numeric(),
    PValue = numeric(),
    Sig = logical()
  )
  
  for (r in response) {
    for (predictor in predictors) {
      if (!(r %in% names(data)) || !(predictor %in% names(data))) {
        next
      }
      cor_test <- cor.test(data[[r]], data[[predictor]], method = "spearman", exact = FALSE)
      
      corr_df <- rbind(corr_df, data.frame(
        Predictor = predictor,
        Response = r,
        Correlation = round(cor_test$estimate, 3),
        PValue = round(cor_test$p.value, 3),
        Sig = ifelse(cor_test$p.value <= 0.05, TRUE, FALSE)
      ))
    }
  }
  
  corr_df <- corr_df |> arrange(desc(Correlation))
  return(corr_df)
}

generate_corr_plot2 <- function(df_names) {
  for (df_name in df_names) {
    
    df <- get(df_name, envir = .GlobalEnv)
    corr <- round(cor(df, method = "spearman", use = "pairwise.complete.obs"), 2)
    matrix <- cor_pmat(df)
    corrplot <- ggcorrplot(corr, p.mat = matrix, method = "square", type = "lower",
                           lab = TRUE, lab_size = 1.2, insig = "blank", title = paste("Correlation Plot:", str_to_title(gsub("\\.", " ", df_name)))) +
      theme(axis.text.x = element_text(size = 4, angle = 90, hjust = 1, vjust = -0.5),
            axis.text.y = element_text(size = 4))
    
    assign(paste(df_name, "_corrplot", sep = ""), corrplot, envir = .GlobalEnv)
  }
}

migrant_health_centers()
state_facility_load()
healthcare()
stateFacilityCorrelations()
```
## What is the relationship between health risk concentrations and healthcare market supply?

#### (Goal is to ultimately assess health care supply-side institutional capacities and possible gaps)

##### First pass: calculate correlations between environmental risk variables and health care services/infrastructure (particularly for data series identified as targeting “migrant” health) but others too



### Exploration


Counts of health facilities labeled as migrant health centers, according to the [National Center for Farmworker Health](https://public.tableau.com/app/profile/national.center.for.farmworker.health/viz/MigrantHealthCentersUDS2023/Dashboard1) (2023). 


An option to export the data to a .csv was not found, so a data set was manually created by investigating each point within the HICAHS states. During the creation of the .csv used in my analysis, some facilities were found to be permanently closed. They have subsequently been excluded from the file.


It must be kept in mind that some facilities in this list are mobile vans or seasonal locations, and three are purely administrative buildings.


Given current knowledge, Wyoming and South Dakota have no facilities labeled as migrant health centers.
$\\$

```{r}
result <- mhc_count |> 
  group_by(state) |> 
  summarise(total = sum(MigrantHealthCenters, na.rm = TRUE))

result <- rbind(result, c("South Dakota", 0))

ggplot(data = result, aes(x = state, y = as.numeric(total), fill = state)) + 
  geom_bar(stat = "identity", color = "black") + 
  labs(x = "State", y = "Migrant Health Centers", title = "Migrant Health Centers Per State (NCFH)") + 
  scale_y_continuous(breaks = seq(from = 0, to = 45, by = 2)) + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

```
```{r, include = FALSE, message = FALSE}
hicahs_States <- invisible(st_read("~/internship/workspace/tl_2024_us_state/tl_2024_us_state.shp"))
```

```{r, echo = FALSE, eval = TRUE}

migrantHealthCenter <- read_csv("~/internship/workspace/migrant_health_centers_ncfh.csv")

migrantHealthCenter <- migrantHealthCenter |> 
  separate(geolocation, into = c("latitude", "longitude"), sep = ",", convert = TRUE)

selected_states <- hicahs_States[hicahs_States$NAME %in% c("Wyoming", "Colorado", "Montana", "Utah", "North Dakota", "South Dakota"), ]
map <- leaflet(migrantHealthCenter) |> 
  addTiles() |> 
  setView(lng = -104.001838, lat = 42.073804, zoom = 5.45) |> 
  addProviderTiles(providers$Esri.WorldTopoMap) |> 
  addCircleMarkers(lng = ~longitude, lat = ~latitude,
                   fill = TRUE, 
                   fillColor = "blue", 
                   fillOpacity = 1,
                   stroke = TRUE, 
                   color = "turquoise", 
                   radius = 3.5) |> 
  addPolygons(data = selected_states,
              fill = FALSE,
              color = "black",
              weight = 0.5,
              opacity = 1)

map
```



### Correlations Between the Total Number of H-2A workers and Healthcare Facilities.
$\\$

Each state has different sets of health facility data, where some states have more or less recorded types of facilities. Utah especially has many more categories of healthcare facilities than any other state. I could remove those differing variables, but then the only facility type in common would be hospitals, and I don't want to limit the data to a singule type. When calculating correlations, I separated data by state, then aggregated all the data sets after calculating the coefficients. Only one correlation with Total H2A Workers was found to be not significant: Critical Access Hospitals in South Dakota.



```{r}
# correlating between number of H-2A workers and the given health facilities

h2aPopulationCorrelations <- h2aPopulationCorrelations |> mutate(PValue = round(PValue, 2)) |> arrange(desc(Correlation))
h2aPopulationCorrelations$Variable <- gsub("_", "", h2aPopulationCorrelations$Variable)
rownames(h2aPopulationCorrelations) <- NULL
kable(h2aPopulationCorrelations, longtable = TRUE, align =c("l", "l", "l", "c", "c", "c"))

#write.csv(h2aPopulationCorrelations, file = "h2aPopulationCorrelations.csv")
```


### Correlations Between Environmental Factors and Healthcare Facilities.

```{r}
weather_facility_load()
create_response_predictors()

colorado_corr <- healthFacility_envir_correlations(colorado_full, response, predictors) |> na.omit()
utah_corr <- healthFacility_envir_correlations(utah_full, response, predictors) |> na.omit()
wyoming_corr <- healthFacility_envir_correlations(wyoming_full, response, predictors) |> na.omit()
northDakota_corr <- healthFacility_envir_correlations(northDakota_full, response, predictors) |> na.omit()
southDakota_corr <- healthFacility_envir_correlations(southDakota_full, response, predictors) |> na.omit()
montana_corr <- healthFacility_envir_correlations(montana_full, response, predictors) |> na.omit()

rownames(montana_corr) <- rownames(utah_corr) <- rownames(wyoming_corr) <- rownames(colorado_corr) <- rownames(northDakota_corr) <- rownames(southDakota_corr) <- NULL

```
$\\$

##### Colorado:

```{r}
cat("Number of Correlations Statistically Signifcant:", sum(colorado_corr$Sig==TRUE), "\n")
cat("Percentage:", round(sum(colorado_corr$Sig==TRUE)/nrow(colorado_corr),4)*100, "%\n")
cat("Number of Correlations Not Statistically Signifcant:",  sum(colorado_corr$Sig==FALSE), "\n")
cat("Percentage:", round(sum(colorado_corr$Sig==FALSE)/nrow(colorado_corr),4)*100, "%")
kable(colorado_corr)
```
$\\$

##### Wyoming:

```{r}
cat("Number of Correlations Statistically Signifcant:", sum(wyoming_corr$Sig==TRUE), "\n")
cat("Percentage:", round(sum(wyoming_corr$Sig==TRUE)/nrow(wyoming_corr),4)*100, "%\n")
cat("Number of Correlations Not Statistically Signifcant:",  sum(wyoming_corr$Sig==FALSE), "\n")
cat("Percentage:", round(sum(wyoming_corr$Sig==FALSE)/nrow(wyoming_corr),4)*100, "%")
kable(wyoming_corr)
```
$\\$

##### Utah:

```{r}
cat("Number of Correlations Statistically Signifcant:", sum(utah_corr$Sig==TRUE), "\n")
cat("Percentage:", round(sum(utah_corr$Sig==TRUE)/nrow(utah_corr),4)*100, "%\n")
cat("Number of Correlations Not Statistically Signifcant:",  sum(utah_corr$Sig==FALSE), "\n")
cat("Percentage:", round(sum(utah_corr$Sig==FALSE)/nrow(utah_corr),4)*100, "%")
kable(utah_corr)
```
$\\$

##### South Dakota:

```{r}
cat("Number of Correlations Statistically Signifcant:", sum(southDakota_corr$Sig==TRUE), "\n")
cat("Percentage:", round(sum(southDakota_corr$Sig==TRUE)/nrow(southDakota_corr),4)*100, "%\n")
cat("Number of Correlations Not Statistically Signifcant:",  sum(southDakota_corr$Sig==FALSE), "\n")
cat("Percentage:", round(sum(southDakota_corr$Sig==FALSE)/nrow(southDakota_corr),4)*100, "%")
kable(southDakota_corr)
```
$\\$

##### North Dakota:

```{r}
cat("Number of Correlations Statistically Signifcant:", sum(northDakota_corr$Sig==TRUE), "\n")
cat("Percentage:", round(sum(northDakota_corr$Sig==TRUE)/nrow(northDakota_corr),4)*100, "%\n")
cat("Number of Correlations Not Statistically Signifcant:",  sum(northDakota_corr$Sig==FALSE), "\n")
cat("Percentage:", round(sum(northDakota_corr$Sig==FALSE)/nrow(northDakota_corr),4)*100, "%")
kable(northDakota_corr)
```
$\\$

##### Montana:

```{r}
cat("Number of Correlations Statistically Signifcant:", sum(montana_corr$Sig==TRUE), "\n")
cat("Percentage:", round(sum(montana_corr$Sig==TRUE)/nrow(montana_corr),4)*100, "%\n")
cat("Number of Correlations Not Statistically Signifcant:",  sum(montana_corr$Sig==FALSE), "\n")
cat("Percentage:", round(sum(montana_corr$Sig==FALSE)/nrow(montana_corr),4)*100, "%")
kable(montana_corr)
```
