---
title: "Health Risks and Worker Density"
author: "Nathan Seto"
date: "10 February 2025"
output: pdf_document
---

```{r setup, include = FALSE}
rm(list = ls())
library(formatR)
library(knitr)
library(tidyverse)
library(stargazer)
library(ggcorrplot)
opts_chunk$set(echo = FALSE, eval = TRUE, tidy = TRUE, warning = FALSE, message = FALSE)
Risk_H2AWorkers <- read_csv("~/internship/workspace/wildfire_disaster.csv")
```

### (1) What is the relationship between health risk factors and worker densities in the HICAHS area?

##### Goal: Calculate unconditional correlations of worker concentrations (total workers) with heat exposure, wildfire exposure, and any other similar correlations of importance.

I started by observing the shape of the data by plotting the total number of workers against variables of interest. Below is a visual examination of the relationship between total workers and two variables of interest: 'Wildfire Risk Score' and 'Average Temperature'.

```{r eda}
plot1 <- ggplot(data = Risk_H2AWorkers, aes(y = total_workers, x = wildfire_hazard_type_risk_index_score)) +
  geom_point(aes(color = county)) +
  geom_smooth(method='lm', se=FALSE, col = "black", linewidth = 0.5) +
  facet_wrap(.~state, scales = "free") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Wildfire Risk Score and Worker Totals",
       x = "Wildfire Hazard Type Risk Index Score",
       y = "Total H2A Workers")

plot2 <- ggplot(data = Risk_H2AWorkers, aes(y = total_workers, x = mean_temperature)) +
  geom_point(aes(color = county)) +
  geom_smooth(method='lm', se=FALSE, col = "black", linewidth = 0.5) +
  facet_wrap(.~state, scales = "free") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Average Temperature and Worker Totals",
       x = "Mean Temperature",
       y = "Total H2A Workers")

plot1
plot2
```

For the rest of my work, I also focused on Wildfire Hazard Type Risk Index Score along with the previous variables.

Following my graphical examination, I calculated correlations using three methods in R. The first method directly calculates only the [correlation](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/cor) between two variables. The results can be found below. The second method calculates the [correlation between a pair of variables and the significance level](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/cor.test). The last creates a [linear model](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/lm) with the response variable being total workers, from which I extracted the $R^2$ value.

Linear Model Summaries:

```{r cor}
MaxTemp_WorkerTotal_lm <- lm(total_workers ~ mean_temperature, data = Risk_H2AWorkers)
FireHzrdIdxScore_WorkerTotal_lm <- lm(total_workers ~ `wildfire_hazard_type_risk_index_score`, data = Risk_H2AWorkers)
FireExpTotal_WorkerTotal_lm <- lm(total_workers ~ `wildfire_exposure_total`, data = Risk_H2AWorkers)
 
cat("Linear Model of Average Temperature and Worker Totals (All States): \n")
stargazer(FireExpTotal_WorkerTotal_lm, type = "text", title = "Summary Statistics of Coefficients")
cat("\nLinear Model of Wild Fire Hazard Index Score and Worker Totals (All States): \n")
stargazer(FireHzrdIdxScore_WorkerTotal_lm, type = "text", title = "Summary Statistics of Coefficients")
cat("\nLinear Model of Fire Exposure Total and Worker Totals (All States): \n")
stargazer(FireExpTotal_WorkerTotal_lm, type = "text", title = "Summary Statistics of Coefficients")
```

Correlation Coefficients:

```{r}
analyze_states <- function(data, states) {
  results <- data.frame()
  
  for (state in states) {
    state_data <- data[data$state == state, ]
    
    predictors <- c("mean_temperature", "wildfire_hazard_type_risk_index_score", "wildfire_exposure_total")
    
    for (predictor in predictors) {
      lm_model <- lm(total_workers ~ get(predictor), data = state_data)
      model_rsq <- summary(lm_model)$r.squared
      
      correlation <- cor(state_data[[predictor]], state_data$total_workers, method = "spearman")
      cor_test <- cor.test(state_data[[predictor]], state_data$total_workers, method = "spearman")
      
      results <- rbind(results, data.frame(
        State = state,
        Predictor = predictor,
        Model_R_sq = model_rsq,
        Spearman_Corr = correlation,
        P_Value = cor_test$p.value
      ))
    }
  }
  
  return(results)
}

states <- c("Colorado", "Montana", "Wyoming", "Utah", "North Dakota", "South Dakota")
results_df <- analyze_states(Risk_H2AWorkers, states)
kable(results_df, caption = "Correlation between Total Workers and Predictors by State", digits = 4)
```

Only a few of the relationships were found to be significantly correlated.

```{r}
cat("Number of Significant Correlations (p < 0.05):", sum(results_df$P_Value<0.05), "\n")
kable(results_df[results_df$P_Value < 0.05, ], caption = "Significant Correlations Found (p < 0.05)", digits = 4)
```

```{r corrplot}
correlation_vector <- c()
  Risk_H2AWorkers_num <- Risk_H2AWorkers[, sapply(Risk_H2AWorkers, is.numeric)]
    N <- colnames(Risk_H2AWorkers_num)
      for (i in N) {
        correlation_value <- cor(x = Risk_H2AWorkers_num[[i]], y = Risk_H2AWorkers_num$total_workers, method = "spearman")
        correlation_vector <- c(correlation_vector, correlation_value)
      }
        correlation_df <- data.frame(column = N, correlation = correlation_vector) |> arrange(desc(correlation))
          correlation_df <-  correlation_df[-1, ]

# Correlogram ggcorrplot()
temp <- Risk_H2AWorkers[sapply(Risk_H2AWorkers, is.numeric)]
corr <- round(cor(temp, method = "spearman"),2)
matrix <- cor_pmat(temp)
corrplot <- ggcorrplot(corr, p.mat = matrix, type = "full",
    lab = TRUE, lab_size = 1.6, insig = "blank", title = "Correlation Plot") +
    theme(axis.text.x = element_text(size = 4.1, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 4.1)
)

cat("\n\n Correlation of all Variables with Total Workers \n")
print(tibble(correlation_df), n = nrow(correlation_df))
cat("\n\n")
corrplot
```
