---
title: "Healthcare"
author: "Nathan Seto"
output: pdf_document
date: "`r Sys.Date()`"
---
```{r setup, include = FALSE}
# rm(list = ls())
library(knitr)
library(formatR)
library(tidyverse)
library(readxl)
library(readr)
library(ggcorrplot)
library(ggplot2)
opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
```

### (3) What is the relationship between health risk concentrations and healthcare market supply?

#### (goal is to ultimately assess health care supply-side institutional capacities and possible gaps)

##### First pass: calculate correlations between environmental risk variables and health care services/infrastructure (particularly for data series identified as targeting “migrant” health) but others too

Data Required:

-   healthcare facilities $\checkmark$

-   immigrant health centers

-   environmental risks $\checkmark$

-   health problems $\checkmark$

-   H2-A population $\checkmark$


Correlations may be less reliable because I cannot group by county because R won't have enough observations to work with. Also, the states have different sets of health facility data, where some states have more/less recorded types of facilities. I could remove those differing variables, but then the only facility type in common would be hospitals, and I don't want to limit the data in this manner. When calculating correlations, I separated data by state, then aggregated all the data sets after calculating the coefficients. Only one correlation with Total H2A Workers was found to be not significant: Critical Access Hospitals in South Dakota.


```{r}
source("C:/Users/natha/OneDrive/Documents/internship/workspace/Processing/Healthcare.R")
state_facility_load()
healthcare()
```


Below is a table of correlations between the total number of H-2A workers and health facilities per county.

```{r}
# correlating between number of H-2A workers and the given health facilities
stateFacilityCorrelations()
kable(h2aPopulationCorrelations)
```

