---
title: "Healthcare"
author: "Nathan Seto"
output: pdf_document
date: "`r Sys.Date()`"
---
```{r setup, include = FALSE}
# rm(list = ls())
library(knitr)
library(formatR)
library(tidyverse)
library(readxl)
library(readr)
library(ggcorrplot)
library(ggplot2)
library(reporter)
opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)

```

### (3) What is the relationship between health risk concentrations and healthcare market supply?

#### (goal is to ultimately assess health care supply-side institutional capacities and possible gaps)

##### First pass: calculate correlations between environmental risk variables and health care services/infrastructure (particularly for data series identified as targeting “migrant” health) but others too

Data Required:

-   healthcare facilities $\checkmark$

-   immigrant health centers

-   environmental risks $\checkmark$

-   health problems $\checkmark$

-   H2-A population $\checkmark$


Correlations may be less reliable because I cannot group by county because R won't have enough observations to work with. Also, the states have different sets of health facility data, where some states have more/less recorded types of facilities. I could remove those differing variables, but then the only facility type in common would be hospitals, and I don't want to limit the data in this manner. When calculating correlations, I separated data by state, then aggregated all the data sets after calculating the coefficients. Only one correlation with Total H2A Workers was found to be not significant: Critical Access Hospitals in South Dakota.


```{r}
source("C:/Users/natha/OneDrive/Documents/internship/workspace/Processing/Healthcare.R")
state_facility_load()
healthcare()
```


Correlations between the total number of H-2A workers and healthcare facilities.

```{r}
# correlating between number of H-2A workers and the given health facilities
stateFacilityCorrelations()
kable(h2aPopulationCorrelations, longtable = TRUE)

# write.csv(h2aPopulationCorrelations, file = "h2aPopulationCorrelations.csv")
```


Correlations between environmental factors and healthcare facilities.

```{r}
weather_facility_load()
create_response_predictors()

colorado_corr <- healthFacility_envir_correlations(colorado_full, response, predictors) |> na.omit()
utah_corr <- healthFacility_envir_correlations(utah_full, response, predictors) |> na.omit()
wyoming_corr <- healthFacility_envir_correlations(wyoming_full, response, predictors) |> na.omit()
northDakota_corr <- healthFacility_envir_correlations(northDakota_full, response, predictors) |> na.omit()
southDakota_corr <- healthFacility_envir_correlations(southDakota_full, response, predictors) |> na.omit()
montana_corr <- healthFacility_envir_correlations(montana_full, response, predictors) |> na.omit()

kable(colorado_corr, longtable = TRUE)
kable(wyoming_corr, longtable = TRUE)
kable(utah_corr, longtable = TRUE)
kable(montana_corr,longtable = TRUE)
kable(southDakota_corr,longtable = TRUE)
kable(northDakota_corr,longtable = TRUE)
# states <- list(
#   colorado = colorado_corr,
#   utah = utah_corr,
#   wyoming = wyoming_corr,
#   northDakota = northDakota_corr,
#   southDakota = southDakota_corr,
#   montana = montana_corr
# )
# 
# 
# for (state in names(states)) {
#   tbl <- create_table(states[[state]]) |>
#     titles(paste(toupper(substr(state, 1, 2)), "Env. and Health Correlation")) |>
#     column_defaults(width = 1, align = "center") |> 
#     define("Predictor", width = 1, align = "left") |>  
#     define("Response", width = 1, align = "left") |>  
#     define("Correlation", width = 0.75, align = "center") |>
#     define("PValue", width = 0.5, align = "center") |>
#     define("Sig", width = 0.5, align = "center")
#   
#   rpt <- create_report(paste("Healthcare_Correlations/", state, ".pdf", sep = ""),
#                        output_type = "PDF", font_size = 10) |>
#     page_footer("Page [pg] of [tpg]") |>
#     add_content(tbl)
#   
#   # Write the report to a file
#   write_report(rpt)
# }
```

