---
title: "Healthcare"
author: "Nathan Seto"
output: pdf_document
date: "`r Sys.Date()`"
---
```{r setup, include = FALSE}
# rm(list = ls())
library(knitr)
library(formatR)
library(tidyverse)
library(readxl)
library(readr)
library(ggcorrplot)
library(ggplot2)
opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
```

### (3) What is the relationship between health risk concentrations and healthcare market supply?

#### (goal is to ultimately assess health care supply-side institutional capacities and possible gaps)

##### First pass: calculate correlations between environmental risk variables and health care services/infrastructure (particularly for data series identified as targeting “migrant” health) but others too

Data Required:

-   healthcare facilities $\checkmark$

-   immigrant health centers

-   environmental risks

-   health problems $\checkmark$

-   H2-A population $\checkmark$


Correlations may be less reliable because I cannot group by county because R won't have enough observations to work with. Also, the states have different sets of health facility data, where some states have more/less recorded types of facilities. I could remove those differing variables, but then the only facility type in common would be hospitals, and I don't want to limit the data in this manner. When calculating correlations, I separated data by state, then aggregated all the data sets after calculating the coefficients. Only one correlation with Total H2A Workers was found to be not significant: Critical Access Hospitals in South Dakota.


```{r}
# health risks
health <- read_csv("~/internship/workspace/PLACES__Local_Data_for_Better_Health__County_Data_2024_release_20250306.csv")
colnames(health) <- tolower(names(health))
state_names <- c("Colorado", "Montana", "North Dakota", "South Dakota", "Utah", "Wyoming")
health <- filter(health, state %in% state_names) |> arrange(stateabbr, county) |> select(-locationid)

# health facilities
colorado <- read_excel("~/internship/workspace/Health Facility Data/ColoradoHealth24_xlsx.xlsx", sheet = 2)
wyoming <- read_excel("~/internship/workspace/Health Facility Data/WyomingHealth24_xlsx.xlsx", sheet = 2)
south_dakota <- read_excel("~/internship/workspace/Health Facility Data/SouthDakotaHealth24_xlsx.xlsx", sheet = 2)
north_dakota <- read_excel("~/internship/workspace/Health Facility Data/NorthDakotaHealth24_xlsx.xlsx")
utah <- read_excel("~/internship/workspace/Health Facility Data/UtahHealth24_xlsx.xlsx", sheet = 2)
montana <- read_excel("~/internship/workspace/Health Facility Data/MontanaHealth24_xlsx.xlsx")

# H2-A population
h2a_population <- read_csv("~/internship/workspace/Written Datasets/h2a_by_county_new.csv")

source("C:/Users/natha/OneDrive/Documents/internship/workspace/Processing/Healthcare.R")
healthcare()

rm("colorado", "montana", "utah", "north_dakota", "south_dakota", "wyoming")
```


```{r}
ggplot(data = na.omit(co_data), aes(y = total_workers_h2a, x = Hospitals, group = county, color = county)) + 
  geom_point(size = 4, alpha = 0.5) + 
  theme_minimal() +
  theme(legend.position = "none") + 
  labs(title = "Hospitals")

```

```{r}
stateFacilityCorrelations()
kable(h2aPopulationCorrelations)
```

