---
title: "H-2A Population, Healthcare, and Envrionmental Risk"
author: "Nathan Seto"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(ggcorrplot)
library(leaflet)
library(leaflet.extras)
library(magrittr)
library(sf)
library(knitr)

opts_chunk$set(echo = F, eval = T, warning = F, message = F)
```

```{r data}
# Data:
H2A <- read_csv("~/internship/workspace/Written Datasets/h2a_by_county_new.csv") # Worker density
MHC <- read_csv("~/internship/workspace/migrant_health_centers_ncfh.csv") # Migrant health centers
ENV <- read_csv("~/internship/workspace/Written Datasets/disaster_cut_clean.csv") # Natural disaster and weather risk
AGO <- read_csv("~/internship/workspace/Written Datasets/ag_output_clean.csv") # Agricultural output
JUL23 <- read_csv("~/internship/HICAHS/Data/Heat_Ag_HumanRisk/CountyMaxTemp_JUL23.csv") # Maximum temperature data for July 2023
AUG23 <- read_csv("~/internship/HICAHS/Data/Heat_Ag_HumanRisk/CountyMaxTemp_AUG23.csv") # Maximum temperature data for August 2023

# Healthcare facilities in each state
# (Cannot bind due to differing number of columns)
COH <- read_excel("~/internship/workspace/Health Facility Data/ColoradoHealth24_xlsx.xlsx", sheet = 2)
WYH <- read_excel("~/internship/workspace/Health Facility Data/WyomingHealth24_xlsx.xlsx", sheet = 2)
SDH <- read_excel("~/internship/workspace/Health Facility Data/SouthDakotaHealth24_xlsx.xlsx", sheet = 2)
NDH <- read_excel("~/internship/workspace/Health Facility Data/NorthDakotaHealth24_xlsx.xlsx", sheet = 1)
UTH <- read_excel("~/internship/workspace/Health Facility Data/UtahHealth24_xlsx.xlsx", sheet = 2)
MTH <- read_excel("~/internship/workspace/Health Facility Data/MontanaHealth24_xlsx.xlsx", sheet = 1)

response <- character()
predictor <- character()
```

```{r cleaning}
# Cleaning:
ENV$county[ENV$county == "lamoure"] <- "la moure"
ENV <- ENV |> select(-matches("coastal|hurricane|tsunami|volcanic", ignore.case = TRUE))
AGO[] <- lapply(AGO, function(col) gsub(",", "", col))

WYH$county <- gsub(" County", "", WYH$county)
MTH$county <- str_to_title(MTH$county)

COH$state <- str_to_title(COH$state)
WYH$state <- str_to_title(WYH$state)
SDH$state <- str_to_title(SDH$state)
NDH$state <- str_to_title(NDH$state)
UTH$state <- str_to_title(UTH$state)
MTH$state <- str_to_title(MTH$state)

JUL23$Name <- gsub(" County", "", JUL23$Name)
AUG23$Name <- gsub(" County", "", AUG23$Name)

JUL23 <- JUL23 |> rename(Jul2023_MaxTemp = Value)
AUG23 <- AUG23 |> rename(Aug2023_MaxTemp = Value)

JUL23 <- JUL23 |> mutate(year = 2023, month = "July", .before = Name)
AUG23 <- AUG23 |> mutate(year = 2023, month = "August", .before = Name)

JUL23 <- JUL23 |> rename(county = Name) |> rename(state = State) |> rename(`July1901-2000Mean` = `1901-2000 Mean`)
AUG23 <- AUG23 |> rename(county = Name) |> rename(state = State) |> rename(`Aug1901-2000Mean` = `1901-2000 Mean`)

# Column uniformity
H2A$county <- str_to_title(H2A$county)
MHC$county <- str_to_title(MHC$county)
ENV$county <- str_to_title(ENV$county)
AGO$county <- str_to_title(AGO$county)

H2A$state <- str_to_title(H2A$state)
MHC$state <- str_to_title(MHC$state)
ENV$state <- str_to_title(ENV$state)
AGO$state <- str_to_title(AGO$state)
AUG23$state <- str_to_title(AUG23$state)
JUL23$state <- str_to_title(JUL23$state)

# Sum number of migrant health facilities per county
MHC_sum <- MHC |>
  group_by(state, county) |>
  summarise(MigrantHealthCenters = n(), .groups = "drop")
```

```{r slicing}
# Binding all data together into one data set to work with:
JOIN <- left_join(ENV, H2A, by = c("state", "county")) # Join H-2A worker totals and natural disaster data sets
JOIN <- left_join(JOIN, AGO, by = c("state", "county")) # Join agricultural output data set
JOIN <- left_join(JOIN, MHC_sum, by = c("state", "county")) # Join migrant health center counts
JOIN <- JOIN |>
  left_join(COH, by = c("state", "county")) |>
  left_join(WYH, by = c("state", "county")) |>
  left_join(SDH, by = c("state", "county")) |>
  left_join(NDH, by = c("state", "county")) |>
  left_join(UTH, by = c("state", "county")) |>
  left_join(MTH, by = c("state", "county"))

JOIN <- JOIN |> mutate(MigrantHealthCenters = ifelse(is.na(MigrantHealthCenters), 0, MigrantHealthCenters))
JOIN <- JOIN |> select(-grep("abbrev", colnames(JOIN), ignore.case = T))

JOIN <- JOIN |> mutate(Hospitals = coalesce(Hospitals.x, Hospitals.y, Hospitals.x.x, Hospitals.y.y, Hospitals.x.x.x, Hospitals.y.y.y))
JOIN <- JOIN |> select(-Hospitals.x, -Hospitals.y, -Hospitals.x.x, -Hospitals.y.y, -Hospitals.x.x.x, -Hospitals.y.y.y)

JOIN <- JOIN |> mutate(`Rural_Health_Clinics` = coalesce(Rural_Health_Clinics.x, Rural_Health_Clinics.y))
JOIN <- JOIN |> select(-Rural_Health_Clinics.x, -Rural_Health_Clinics.y)
#---------------------------------------------------------------------------------------------------------------#
# Subsetting
JOIN_CUT <- JOIN |>
  select(
    "state",
    "county",
    "Population2020",
    "BuildingValue",
    "AgricultureValue",
    "Areasqmi",
    "H2A_workers",
    grep(
      "income|sales|farm_sales|income_net_|commodity_totals|crop_totals_sales|drought|wildfire|heat|heatwave",
      colnames(JOIN),
      ignore.case = TRUE
    ),
    349:length(JOIN)
  )

# Very simplified subset of risk variables
JOIN_CUT <- JOIN_CUT |> select(-grep("Expected|wheat|cattle|hogs|chickens|receipts|historic|index", colnames(JOIN_CUT), ignore.case = T))
# Filling NAs
JOIN_CUT[sapply(JOIN_CUT, is.numeric)] <- lapply(JOIN_CUT[sapply(JOIN_CUT, is.numeric)], function(x) ifelse(is.na(x), 0, x))
# Deleting empty columns
JOIN_CUT <- JOIN_CUT[, colSums(!is.na(JOIN_CUT)) > 0]

JOIN_CUT <- JOIN_CUT |> left_join(JUL23 |>
                                    select(state, county, Jul2023_MaxTemp, `July1901-2000Mean`), by = c("state", "county")) |>
  left_join(AUG23 |>
              select(state, county, Aug2023_MaxTemp, `Aug1901-2000Mean`), by = c("state", "county"))
```


## Correlation between H-2A population & Wildfire/Drought/HeatWave Exposure, Jul/Aug Max Temp & 99-year Mean Temp

```{r}
BURN <- JOIN_CUT |> select("H2A_workers", matches("drought|wildfire|heat|heatwave|Aug|Jul|July"))
BURN_CORR <- data.frame(Variable = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (predictor in colnames(BURN)) {
  test <- cor.test(BURN$H2A_workers, BURN[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
  BURN_CORR <- rbind(BURN_CORR, data.frame(Variable = predictor,
                                           Correlation = round(test$estimate, 3),
                                           PValue = round(test$p.value,3 )))
}

BURN_CORR <- BURN_CORR |> arrange(desc(Correlation))
BURN_CORR$Significance <- ifelse(BURN_CORR$PValue <= 0.05, TRUE, FALSE)
BURN_CORR <- BURN_CORR[-1,]
rownames(BURN_CORR) <- NULL
kable(BURN_CORR, caption = "Correlation Between H-2A Population and Fire/Heat Environmental Risk")

# Corrplot
COR_MTX <- round(cor(BURN, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "black",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 1.7,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  )
```


## Correlation between H-2A population & migrant healthcare centers

```{r}
H2AMHC <- H2A |> left_join(MHC_sum, by = c("state", "county"))
H2AMHC$MigrantHealthCenters[is.na(H2AMHC$MigrantHealthCenters)] <- 0

MHC_CORR <- data.frame(Variable = character(),
                       Correlation = numeric(),
                       PValue = numeric() )

test <- cor.test(H2AMHC$H2A_workers, H2AMHC$MigrantHealthCenters, method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
MHC_CORR <- rbind(MHC_CORR, data.frame(Variable = "MigrantHealthCenters",
                                       Correlation = round(test$estimate, 3),
                                       PValue = round(test$p.value, 3 )))

MHC_CORR$Significance <- ifelse(MHC_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(MHC_CORR) <- NULL
kable(MHC_CORR, caption = "Correlation between H-2A Population and Migrant Health Centers")
# Low Correlation: Gap in healthcare for migrant workers (statistically significant)
```


## Correlation between healthcare facilities per capita, migrant healthcare centers (not adjusted) & Wildfire/Drought/Heat Wave Exposure

```{r}
FIRE <- JOIN_CUT |> select(40:length(JOIN_CUT), matches("drought|wildfire|heat|heatwave"))
colnames(FIRE) <- gsub(" ", "", colnames(FIRE))

FIRE_HC <- FIRE |> select(1:20)
FIRE_HC <- cbind(FIRE_HC, Population2020 = JOIN$Population2020)
FIRE_HC <- FIRE_HC |>
  mutate(across(
    .cols = -c(MigrantHealthCenters, Population2020),
    .fns = ~ .x / Population2020
  ))
colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")] <-
  paste0(colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")], "PerCapita")
FIRE_HC <- FIRE_HC |> select(-Population2020)

FIRE_PRED <- FIRE |> select(21:length(FIRE))

FIRE_CORR <- data.frame(Response = character(),
                        Predictor = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (response in colnames(FIRE_HC)) {
  for( predictor in colnames(FIRE_PRED)) {
    FIRE_test <- cor.test(FIRE_HC[[response]], FIRE_PRED[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
    FIRE_CORR <- rbind(FIRE_CORR, data.frame(Response = response, Predictor = predictor, Correlation = round(FIRE_test$estimate, 3), PValue = round(FIRE_test$p.value, 3)))
  }
}

FIRE_CORR <- FIRE_CORR |> arrange(desc(Correlation))
FIRE_CORR$Significance <- ifelse(FIRE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(FIRE_CORR) <- NULL
kable(FIRE_CORR, caption = "Correlation between healthcare facilities and Fire/Heat Environmental Risk")
```



## Correlation between farm sales & Wildfire/Drought/HeatWave Exposure

```{r, fig.width = 8, fig.height = 10}
BLAZE_PRED <- JOIN_CUT |> select(matches("drought|wildfire|heat|heatwave"))
BLAZE_AGO <- JOIN_CUT |> select(intersect(colnames(AGO), colnames(JOIN_CUT))) |> select(-state, -county)
BLAZE_AGO[] <- lapply(BLAZE_AGO, as.numeric)
BLAZE_AGO[is.na(BLAZE_AGO)] <- 0
BLAZE <- cbind(BLAZE_PRED, BLAZE_AGO)

colnames(BLAZE) <- gsub("_", " ", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE), perl = TRUE)
colnames(BLAZE) <- gsub(" ", "", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\$", "Dollars", colnames(BLAZE))

colnames(BLAZE_AGO) <- gsub("_", " ", colnames(BLAZE_AGO))
colnames(BLAZE_AGO) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE_AGO), perl = TRUE)
colnames(BLAZE_AGO) <- gsub(" ", "", colnames(BLAZE_AGO))
colnames(BLAZE_AGO) <- gsub("\\$", "Dollars", colnames(BLAZE_AGO))

colnames(BLAZE_PRED) <- gsub("_", " ", colnames(BLAZE_PRED))
colnames(BLAZE_PRED) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE_PRED), perl = TRUE)
colnames(BLAZE_PRED) <- gsub(" ", "", colnames(BLAZE_PRED))
colnames(BLAZE_PRED) <- gsub("\\$", "Dollars", colnames(BLAZE_PRED))


response <- character()
predictor <- character()

BLAZE_CORR <- data.frame(Predictor = character(),
                        Response = character(),
                        Correlation = numeric(),
                        PValue = numeric())

for (response in colnames(BLAZE_AGO)) {
  for (predictor in colnames(BLAZE_PRED)) {
      BLAZE_test <- cor.test(BLAZE[[response]], BLAZE[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
      BLAZE_CORR <- rbind(BLAZE_CORR, data.frame(Predictor = predictor, Response = response, Correlation = round(BLAZE_test$estimate, 3), PValue = round(BLAZE_test$p.value, 3)))
  }
}

BLAZE_CORR <- BLAZE_CORR |> arrange(desc(Correlation))
BLAZE_CORR$Significance <- ifelse(BLAZE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(BLAZE_CORR) <- NULL
kable(BLAZE_CORR, caption = "Correlation between Farm output and Wildfire/Drought/HeatWave Exposure")

# Corrplot
COR_MTX <- round(cor(BLAZE, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "black",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 1.34,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  )
```



### Appendix: 
```{r appendix, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(ggcorrplot)
library(leaflet)
library(leaflet.extras)
library(magrittr)
library(sf)
#---------------------------------------------------------------------------------------------------------------#
# Data:
H2A <- read_csv("filepath/h2a_by_county_new.csv") # Worker density
MHC <- read_csv("filepath/migrant_health_centers_ncfh.csv") # Migrant health centers
ENV <- read_csv("filepath/disaster_cut_clean.csv") # Natural disaster and weather risk
AGO <- read_csv("filepath/ag_output_clean.csv") # Agricultural output
JUL23 <- read_csv("filepath/CountyMaxTemp_JUL23.csv") # Maximum temperature data for July 2023
AUG23 <- read_csv("filepath/CountyMaxTemp_AUG23.csv") # Maximum temperature data for August 2023

# Healthcare facilities in each state
# (Cannot bind due to differing number of columns)
COH <- read_excel("filepath/ColoradoHealth24_xlsx.xlsx", sheet = 2)
WYH <- read_excel("filepath/WyomingHealth24_xlsx.xlsx", sheet = 2)
SDH <- read_excel("filepath/SouthDakotaHealth24_xlsx.xlsx", sheet = 2)
NDH <- read_excel("filepath/NorthDakotaHealth24_xlsx.xlsx", sheet = 1)
UTH <- read_excel("filepath/UtahHealth24_xlsx.xlsx", sheet = 2)
MTH <- read_excel("filepath/MontanaHealth24_xlsx.xlsx", sheet = 1)

response <- character()
predictor <- character()
#---------------------------------------------------------------------------------------------------------------#
# Cleaning:
ENV$county[ENV$county == "lamoure"] <- "la moure"
ENV <- ENV |> select(-matches("coastal|hurricane|tsunami|volcanic", ignore.case = TRUE))
AGO[] <- lapply(AGO, function(col) gsub(",", "", col))

WYH$county <- gsub(" County", "", WYH$county)
MTH$county <- str_to_title(MTH$county)

COH$state <- str_to_title(COH$state)
WYH$state <- str_to_title(WYH$state)
SDH$state <- str_to_title(SDH$state)
NDH$state <- str_to_title(NDH$state)
UTH$state <- str_to_title(UTH$state)
MTH$state <- str_to_title(MTH$state)

JUL23$Name <- gsub(" County", "", JUL23$Name)
AUG23$Name <- gsub(" County", "", AUG23$Name)

JUL23 <- JUL23 |> rename(Jul2023_MaxTemp = Value)
AUG23 <- AUG23 |> rename(Aug2023_MaxTemp = Value)

JUL23 <- JUL23 |> mutate(year = 2023, month = "July", .before = Name)
AUG23 <- AUG23 |> mutate(year = 2023, month = "August", .before = Name)

JUL23 <- JUL23 |> rename(county = Name) |> rename(state = State) |> rename(`July1901-2000Mean` = `1901-2000 Mean`)
AUG23 <- AUG23 |> rename(county = Name) |> rename(state = State) |> rename(`Aug1901-2000Mean` = `1901-2000 Mean`)

# Column uniformity
H2A$county <- str_to_title(H2A$county)
MHC$county <- str_to_title(MHC$county)
ENV$county <- str_to_title(ENV$county)
AGO$county <- str_to_title(AGO$county)

H2A$state <- str_to_title(H2A$state)
MHC$state <- str_to_title(MHC$state)
ENV$state <- str_to_title(ENV$state)
AGO$state <- str_to_title(AGO$state)
AUG23$state <- str_to_title(AUG23$state)
JUL23$state <- str_to_title(JUL23$state)

# Sum number of migrant health facilities per county
MHC_sum <- MHC |>
  group_by(state, county) |>
  summarise(MigrantHealthCenters = n(), .groups = "drop")
#---------------------------------------------------------------------------------------------------------------#
# Binding all data together into one data set to work with:
JOIN <- left_join(ENV, H2A, by = c("state", "county")) # Join H-2A worker totals and natural disaster data sets
JOIN <- left_join(JOIN, AGO, by = c("state", "county")) # Join agricultural output data set
JOIN <- left_join(JOIN, MHC_sum, by = c("state", "county")) # Join migrant health center counts
JOIN <- JOIN |>
  left_join(COH, by = c("state", "county")) |>
  left_join(WYH, by = c("state", "county")) |>
  left_join(SDH, by = c("state", "county")) |>
  left_join(NDH, by = c("state", "county")) |>
  left_join(UTH, by = c("state", "county")) |>
  left_join(MTH, by = c("state", "county"))

JOIN <- JOIN |> mutate(MigrantHealthCenters = ifelse(is.na(MigrantHealthCenters), 0, MigrantHealthCenters))
JOIN <- JOIN |> select(-grep("abbrev", colnames(JOIN), ignore.case = T))

JOIN <- JOIN |> mutate(Hospitals = coalesce(Hospitals.x, Hospitals.y, Hospitals.x.x, Hospitals.y.y, Hospitals.x.x.x, Hospitals.y.y.y))
JOIN <- JOIN |> select(-Hospitals.x, -Hospitals.y, -Hospitals.x.x, -Hospitals.y.y, -Hospitals.x.x.x, -Hospitals.y.y.y)

JOIN <- JOIN |> mutate(`Rural_Health_Clinics` = coalesce(Rural_Health_Clinics.x, Rural_Health_Clinics.y))
JOIN <- JOIN |> select(-Rural_Health_Clinics.x, -Rural_Health_Clinics.y)
#---------------------------------------------------------------------------------------------------------------#
# Subsetting
JOIN_CUT <- JOIN |>
  select(
    "state",
    "county",
    "Population2020",
    "BuildingValue",
    "AgricultureValue",
    "Areasqmi",
    "H2A_workers",
    grep(
      "income|sales|farm_sales|income_net_|commodity_totals|crop_totals_sales|drought|wildfire|heat|heatwave",
      colnames(JOIN),
      ignore.case = TRUE
    ),
    349:length(JOIN)
  )

# Very simplified subset of risk variables
JOIN_CUT <- JOIN_CUT |> select(-grep("Expected|wheat|cattle|hogs|chickens|receipts|historic|index", colnames(JOIN_CUT), ignore.case = T))
# Filling NAs
JOIN_CUT[sapply(JOIN_CUT, is.numeric)] <- lapply(JOIN_CUT[sapply(JOIN_CUT, is.numeric)], function(x) ifelse(is.na(x), 0, x))
# Deleting empty columns
JOIN_CUT <- JOIN_CUT[, colSums(!is.na(JOIN_CUT)) > 0]

JOIN_CUT <- JOIN_CUT |> left_join(JUL23 |>
                                    select(state, county, Jul2023_MaxTemp, `July1901-2000Mean`), by = c("state", "county")) |>
  left_join(AUG23 |>
              select(state, county, Aug2023_MaxTemp, `Aug1901-2000Mean`), by = c("state", "county"))
#---------------------------------------------------------------------------------------------------------------#
# 1) Correlation between H-2A population & Wildfire/Drought/HeatWave Exposure, Jul/Aug Max Temp & 99-year Mean Temp

BURN <- JOIN_CUT |> select("H2A_workers", matches("drought|wildfire|heat|heatwave|Aug|Jul|July"))
BURN_CORR <- data.frame(Variable = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (predictor in colnames(BURN)) {
  test <- cor.test(BURN$H2A_workers, BURN[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
  BURN_CORR <- rbind(BURN_CORR, data.frame(Variable = predictor,
                                           Correlation = round(test$estimate, 3),
                                           PValue = round(test$p.value,3 )))
}

BURN_CORR <- BURN_CORR |> arrange(desc(Correlation))
BURN_CORR$Significance <- ifelse(BURN_CORR$PValue <= 0.05, TRUE, FALSE)
BURN_CORR <- BURN_CORR[-1,]
rownames(BURN_CORR) <- NULL
kable(BURN_CORR, caption = "Correlation Between H-2A Population and Fire/Heat Environmental Risk")

# Corrplot
COR_MTX <- round(cor(BURN, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "black",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 1.7,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  )
#---------------------------------------------------------------------------------------------------------------#
# 2) Correlation between H-2A population & migrant healthcare centers

H2AMHC <- H2A |> left_join(MHC_sum, by = c("state", "county"))
H2AMHC$MigrantHealthCenters[is.na(H2AMHC$MigrantHealthCenters)] <- 0

MHC_CORR <- data.frame(Variable = character(),
                       Correlation = numeric(),
                       PValue = numeric() )

test <- cor.test(H2AMHC$H2A_workers, H2AMHC$MigrantHealthCenters, method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
MHC_CORR <- rbind(MHC_CORR, data.frame(Variable = "MigrantHealthCenters",
                                       Correlation = round(test$estimate, 3),
                                       PValue = round(test$p.value, 3 )))

MHC_CORR$Significance <- ifelse(MHC_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(MHC_CORR) <- NULL
kable(MHC_CORR, caption = "Correlation between H-2A Population and Migrant Health Centers")
# Low Correlation: Gap in healthcare for migrant workers (statistically significant)
#---------------------------------------------------------------------------------------------------------------#
# 3) Correlation between healthcare facilities per capita, migrant healthcare centers (not adjusted) & Wildfire/Drought/Heat Wave Exposure

FIRE <- JOIN_CUT |> select(40:length(JOIN_CUT), matches("drought|wildfire|heat|heatwave"))
colnames(FIRE) <- gsub(" ", "", colnames(FIRE))

FIRE_HC <- FIRE |> select(1:20)
FIRE_HC <- cbind(FIRE_HC, Population2020 = JOIN$Population2020)
FIRE_HC <- FIRE_HC |>
  mutate(across(
    .cols = -c(MigrantHealthCenters, Population2020),
    .fns = ~ .x / Population2020
  ))
colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")] <-
  paste0(colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")], "PerCapita")
FIRE_HC <- FIRE_HC |> select(-Population2020)

FIRE_PRED <- FIRE |> select(21:length(FIRE))

FIRE_CORR <- data.frame(Response = character(),
                        Predictor = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (response in colnames(FIRE_HC)) {
  for( predictor in colnames(FIRE_PRED)) {
    FIRE_test <- cor.test(FIRE_HC[[response]], FIRE_PRED[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
    FIRE_CORR <- rbind(FIRE_CORR, data.frame(Response = response, Predictor = predictor, Correlation = round(FIRE_test$estimate, 3), PValue = round(FIRE_test$p.value, 3)))
  }
}

FIRE_CORR <- FIRE_CORR |> arrange(desc(Correlation))
FIRE_CORR$Significance <- ifelse(FIRE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(FIRE_CORR) <- NULL
kable(FIRE_CORR, caption = "Correlation between healthcare facilities and Fire/Heat Environmental Risk")
#---------------------------------------------------------------------------------------------------------------#
# 4) Correlation between farm sales & Wildfire/Drought/HeatWave Exposure

BLAZE_PRED <- JOIN_CUT |> select(matches("drought|wildfire|heat|heatwave"))
BLAZE_AGO <- JOIN_CUT |> select(intersect(colnames(AGO), colnames(JOIN_CUT))) |> select(-state, -county)
BLAZE_AGO[] <- lapply(BLAZE_AGO, as.numeric)
BLAZE_AGO[is.na(BLAZE_AGO)] <- 0
BLAZE <- cbind(BLAZE_PRED, BLAZE_AGO)

colnames(BLAZE) <- gsub("_", " ", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE), perl = TRUE)
colnames(BLAZE) <- gsub(" ", "", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\$", "Dollars", colnames(BLAZE))


response <- character()
predictor <- character()

BLAZE_CORR <- data.frame(Predictor = character(),
                        Response = character(),
                        Correlation = numeric(),
                        PValue = numeric())

for (response in colnames(BLAZE_AGO)) {
  for (predictor in colnames(BLAZE_PRED)) {
      BLAZE_test <- cor.test(BLAZE_AGO[[response]], BLAZE_PRED[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
      BLAZE_CORR <- rbind(BLAZE_CORR, data.frame(Predictor = predictor, Response = response, Correlation = round(BLAZE_test$estimate, 3), PValue = round(BLAZE_test$p.value, 3)))
  }
}

BLAZE_CORR <- BLAZE_CORR |> arrange(desc(Correlation))
BLAZE_CORR$Significance <- ifelse(BLAZE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(BLAZE_CORR) <- NULL
kable(BLAZE_CORR, caption = "Correlation between Farm output and Wildfire/Drought/HeatWave Exposure")

# Corrplot
COR_MTX <- round(cor(BLAZE, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "black",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 1.7,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  )
```

