---
title: "H-2A Population, Healthcare, and Exposure to Envrionmental Risk"
author: "Nathan Seto"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(ggcorrplot)
library(leaflet)
library(leaflet.extras)
library(magrittr)
library(sf)
library(knitr)

opts_chunk$set(echo = F, eval = T, warning = F, message = F)
```

```{r data, message=FALSE, warning=FALSE}
# Data:
H2A <- read_csv("~/internship/workspace/Written Datasets/h2a_by_county_new.csv") # Worker density
MHC <- read_csv("~/internship/workspace/migrant_health_centers_ncfh.csv") # Migrant health centers
ENV <- read_csv("~/internship/workspace/Written Datasets/disaster_cut_clean.csv") # Natural disaster and weather risk
AGO <- read_csv("~/internship/workspace/Written Datasets/ag_output_clean.csv") # Agricultural output
JUL23 <- read_csv("~/internship/HICAHS/Data/Heat_Ag_HumanRisk/CountyMaxTemp_JUL23.csv") # Maximum temperature data for July 2023
AUG23 <- read_csv("~/internship/HICAHS/Data/Heat_Ag_HumanRisk/CountyMaxTemp_AUG23.csv") # Maximum temperature data for August 2023

# Healthcare facilities in each state
# (Cannot bind due to differing number of columns)
COH <- read_excel("~/internship/workspace/Health Facility Data/ColoradoHealth24_xlsx.xlsx", sheet = 2)
WYH <- read_excel("~/internship/workspace/Health Facility Data/WyomingHealth24_xlsx.xlsx", sheet = 2)
SDH <- read_excel("~/internship/workspace/Health Facility Data/SouthDakotaHealth24_xlsx.xlsx", sheet = 2)
NDH <- read_excel("~/internship/workspace/Health Facility Data/NorthDakotaHealth24_xlsx.xlsx", sheet = 1)
UTH <- read_excel("~/internship/workspace/Health Facility Data/UtahHealth24_xlsx.xlsx", sheet = 2)
MTH <- read_excel("~/internship/workspace/Health Facility Data/MontanaHealth24_xlsx.xlsx", sheet = 1)

# Shape file for mapping
HICAHS_SHP <- invisible(st_read("~/internship/workspace/tl_2024_us_state/tl_2024_us_state.shp", quiet = TRUE))
COUNTY_SHP <- invisible(st_read("~/internship/workspace/tl_2024_us_county/tl_2024_us_county.shp", quiet = TRUE))

response <- character()
predictor <- character()
```

```{r cleaning}
ENV$county[ENV$county == "lamoure"] <- "la moure"
ENV <- ENV |> select(-matches("coastal|hurricane|tsunami|volcanic", ignore.case = TRUE))
AGO[] <- lapply(AGO, function(col) gsub(",", "", col))

WYH$county <- gsub(" County", "", WYH$county)
MTH$county <- str_to_title(MTH$county)

COH$state <- str_to_title(COH$state)
WYH$state <- str_to_title(WYH$state)
SDH$state <- str_to_title(SDH$state)
NDH$state <- str_to_title(NDH$state)
UTH$state <- str_to_title(UTH$state)
MTH$state <- str_to_title(MTH$state)

JUL23$Name <- gsub(" County", "", JUL23$Name)
AUG23$Name <- gsub(" County", "", AUG23$Name)

JUL23 <- JUL23 |> rename(Jul2023_MaxTemp = Value)
AUG23 <- AUG23 |> rename(Aug2023_MaxTemp = Value)

JUL23 <- JUL23 |> mutate(year = 2023, month = "July", .before = Name)
AUG23 <- AUG23 |> mutate(year = 2023, month = "August", .before = Name)

JUL23 <- JUL23 |> rename(county = Name) |> rename(state = State) |> rename(`July1901-2000Mean` = `1901-2000 Mean`)
AUG23 <- AUG23 |> rename(county = Name) |> rename(state = State) |> rename(`Aug1901-2000Mean` = `1901-2000 Mean`)

# Column uniformity
H2A$county <- str_to_title(H2A$county)
MHC$county <- str_to_title(MHC$county)
ENV$county <- str_to_title(ENV$county)
AGO$county <- str_to_title(AGO$county)

H2A$state <- str_to_title(H2A$state)
MHC$state <- str_to_title(MHC$state)
ENV$state <- str_to_title(ENV$state)
AGO$state <- str_to_title(AGO$state)
AUG23$state <- str_to_title(AUG23$state)
JUL23$state <- str_to_title(JUL23$state)

# Adding latitude and longitude for mapping
MHC <- MHC |> separate(geolocation, into = c("latitude", "longitude"), sep = ",", convert = TRUE)

# Sum number of migrant health facilities per county
MHC_sum <- MHC |>
  group_by(state, county) |>
  summarise(MigrantHealthCenters = n(), .groups = "drop")
```

```{r}
# Binding all data together into one data set to work with:
JOIN <- left_join(ENV, H2A, by = c("state", "county")) # Join H-2A worker totals and natural disaster data sets
JOIN <- left_join(JOIN, AGO, by = c("state", "county")) # Join agricultural output data set
JOIN <- left_join(JOIN, MHC_sum, by = c("state", "county")) # Join migrant health center counts
JOIN <- JOIN |>
  left_join(COH, by = c("state", "county")) |>
  left_join(WYH, by = c("state", "county")) |>
  left_join(SDH, by = c("state", "county")) |>
  left_join(NDH, by = c("state", "county")) |>
  left_join(UTH, by = c("state", "county")) |>
  left_join(MTH, by = c("state", "county"))

JOIN <- JOIN |> mutate(MigrantHealthCenters = ifelse(is.na(MigrantHealthCenters), 0, MigrantHealthCenters))
JOIN <- JOIN |> select(-grep("abbrev", colnames(JOIN), ignore.case = T))

JOIN <- JOIN |> mutate(Hospitals = coalesce(Hospitals.x, Hospitals.y, Hospitals.x.x, Hospitals.y.y, Hospitals.x.x.x, Hospitals.y.y.y))
JOIN <- JOIN |> select(-Hospitals.x, -Hospitals.y, -Hospitals.x.x, -Hospitals.y.y, -Hospitals.x.x.x, -Hospitals.y.y.y)

JOIN <- JOIN |> mutate(`RuralHealthClinics` = coalesce(`Rural_Clinics`, `Rural Health Clinics`, Rural_Health_Clinics.x, Rural_Health_Clinics.y))
JOIN <- JOIN |> select(-`Rural_Clinics`, -`Rural Health Clinics`, -Rural_Health_Clinics.x, -Rural_Health_Clinics.y)

JOIN <- JOIN |> mutate(`CriticalAccessHospitals` = coalesce(`Critical Access Hospitals`, Critical_Access_Hospitals))
JOIN <- JOIN |> select(-`Critical Access Hospitals`, -Critical_Access_Hospitals)

JOIN$county[JOIN$county == "lamoure"] <- "la moure"
#---------------------------------------------------------------------------------------------------------------#
JOIN_CUT <- JOIN |>
  select(
    "state",
    "county",
    "Population2020",
    "BuildingValue",
    "AgricultureValue",
    "Areasqmi",
    "H2A_workers",
    "JOIN_CUT$latitude",
    "JOIN_CUT$longitude",
    grep(
      "income|sales|farm_sales|income_net_|commodity_totals|crop_totals_sales|drought|wildfire|heat|heatwave",
      colnames(JOIN),
      ignore.case = TRUE
    ),
    349:length(JOIN)
  )

# Very simplified subset of risk variables
JOIN_CUT <- JOIN_CUT |> select(-grep("Expected|wheat|cattle|hogs|chickens|receipts|historic|index", colnames(JOIN_CUT), ignore.case = T))
# Filling NAs
JOIN_CUT[sapply(JOIN_CUT, is.numeric)] <- lapply(JOIN_CUT[sapply(JOIN_CUT, is.numeric)], function(x) ifelse(is.na(x), 0, x))
# Deleting empty columns
JOIN_CUT <- JOIN_CUT[, colSums(!is.na(JOIN_CUT)) > 0]

JOIN_CUT <- JOIN_CUT |>
  left_join(JUL23 |> select(state, county, Jul2023_MaxTemp, `July1901-2000Mean`), by = c("state", "county")) |>
  left_join(AUG23 |> select(state, county, Aug2023_MaxTemp, `Aug1901-2000Mean`), by = c("state", "county"))

JOIN_CUT <- JOIN_CUT |> arrange(state, county)
```


The goal of this project is to evaluate the correlation between aspects of migrant agricultural worker population and environmental risk within the states of interest to the High Plains Intermountain Center for Agricultural Health and Safety [(HICAHS)](https://vetmedbiosci.colostate.edu/hicahs/). HICAHS is involved with Colorado, Wyoming, Utah, Montana, North Dakota, and South Dakota.


The main goals are as follows: 1) investigate the relationship between health risk concentrations and healthcare market supply, particularly the supply of migrant health centers, 2) investigate the relationship between health risk factors and worker densities in the HICAHS area,  3) investigate the relationship between health risk concentrations and economic output of agriculture.


The Federal Emergency Management Agency (FEMA) defines exposure to some natural disaster as "...the representative value of buildings (in dollars), population (in both people and population equivalence dollars), or agriculture (in dollars) potentially exposed to a natural hazard occurrence." In other words, locations with higher building value, agricultural value, and/or population density would have higher exposure values associated with them. 


A migrant health center is a specialized healthcare facility that provides services to migrant agricultural workers and their families. 


## Exploration


The figures below show the distribution of migrant health centers and H-2A population density in the HICAHS area In total. There are 56 health centers between all six states, with Colorado having the most, and South Dakota and Wyoming having none, and 12,893 H-2A agricultural workers.

```{r}
GGDAT1 <- MHC_sum |> 
  group_by(state) |> 
  summarise(total = sum(MigrantHealthCenters, na.rm = TRUE))

GGDAT1 <- rbind(GGDAT1, c("South Dakota", 0), c("Wyoming", 0))

GGPLOT1 <- ggplot(data = GGDAT1, aes(x = state, y = as.numeric(total))) + 
  geom_bar(stat = "identity", color = "black", fill = "darkgreen", width = .75) + 
  labs(x = "State", y = "Migrant Health Centers", title = "Figure 2: Migrant Health Centers Per State (NCFH 2023)") + 
  scale_y_continuous(breaks = seq(from = 0, to = 40, by = 2)) + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

GGDAT2 <- H2A |> 
  group_by(state) |> 
  summarise(total = sum(H2A_workers, na.rm = TRUE))

GGDAT1$total <- as.numeric(GGDAT1$total)
GGDAT2$total <- as.numeric(GGDAT2$total)

GGDAT1$state <- as.factor(GGDAT1$state)
GGDAT2$state <- as.factor(GGDAT2$state)

GGPLOT2 <- ggplot(data = GGDAT2, aes(x = state, y = as.numeric(total))) + 
  geom_bar(stat = "identity", color = "black", fill = "skyblue", width = .75) + 
  labs(x = "State", y = "Worker Count", title = "Figure 1: H-2A Workers") + 
  scale_y_continuous(breaks = seq(from = 0, to = max(GGDAT2$total), by = 500)) + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

GGPLOT2
GGPLOT1

```


#### Figure 3: Geographic Distribution of H-2A worker population and migrant healthcare centers

```{r}
selected_states <- HICAHS_SHP[HICAHS_SHP$NAME %in% c("Wyoming", "Colorado", "Montana", "Utah", "North Dakota", "South Dakota"), ]
selected_counties <- COUNTY_SHP[COUNTY_SHP$STATEFP %in% selected_states$STATEFP, ] |> arrange(STATEFP, NAME)



JOIN_CUT_CLEAN <- JOIN_CUT |> filter(county != "Lamoure")
JOIN_CUT_sf <- st_as_sf(JOIN_CUT_CLEAN, coords = c("JOIN_CUT$longitude", "JOIN_CUT$latitude"), crs = 4326)


color_palette <- colorNumeric(palette = c("lightblue", "violet", "red"), domain = JOIN_CUT_sf$H2A_workers)
map <- leaflet(MHC) |>
  addTiles() |>
  setView(lng = -104.993498, lat = 42.468594, zoom = 5.45) |>
  addProviderTiles(providers$Esri.WorldTopoMap) |>
  addPolygons(data = selected_states,
              fill = FALSE,
              color = "black",
              weight = 0.5,
              opacity = 1) |>
  addPolygons(data = selected_counties,
              fillColor = ~color_palette(JOIN_CUT_sf$H2A_workers),
              fillOpacity = 0.7,
              color = "darkgrey",
              weight = 0.5,
              opacity = 1) |>
  addCircleMarkers(lng = ~longitude, lat = ~latitude,
                   fill = TRUE,
                   fillColor = "blue",
                   fillOpacity = 1,
                   stroke = T,
                   radius = 2) |>
  addLegend(pal = color_palette, values = JOIN_CUT_sf$H2A_workers, title = "H-2A Worker Population", position = "bottomright")



map
```


## Correlation between H-2A population and migrant healthcare centers


The correlation between the population of workers and migrant health centers is dangerously low (and statistically significant), signaling a vastly outclassed healthcare supply for migrant workers and their families.

```{r}
H2AMHC <- H2A |> left_join(MHC_sum, by = c("state", "county"))
H2AMHC$MigrantHealthCenters[is.na(H2AMHC$MigrantHealthCenters)] <- 0

MHC_CORR <- data.frame(Variable = character(),
                       Correlation = numeric(),
                       PValue = numeric() )

test <- cor.test(H2AMHC$H2A_workers, H2AMHC$MigrantHealthCenters, method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
MHC_CORR <- rbind(MHC_CORR, data.frame(Variable = "MigrantHealthCenters",
                                       Correlation = round(test$estimate, 3),
                                       PValue = round(test$p.value, 3 )))

MHC_CORR$Significance <- ifelse(MHC_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(MHC_CORR) <- NULL
kable(MHC_CORR, caption = "Table 1: Correlation between H-2A Population and Migrant Health Centers")
# Low Correlation: Gap in healthcare for migrant workers (statistically significant)
```


## Correlation between H-2A population and wildfire, drought, and heat wave exposure, and temperature


The correlation coefficients found between H-2A worker density and exposure risk is not surprising. 

```{r, fig.width = 8, fig.height = 10}
BURN <- JOIN_CUT |> select("H2A_workers", matches("drought|wildfire|heat|heatwave|Aug|Jul|July"))
BURN_CORR <- data.frame(Variable = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (predictor in colnames(BURN)) {
  test <- cor.test(BURN$H2A_workers, BURN[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
  BURN_CORR <- rbind(BURN_CORR, data.frame(Variable = predictor,
                                           Correlation = round(test$estimate, 3),
                                           PValue = round(test$p.value,3 )))
}

BURN_CORR <- BURN_CORR |> arrange(desc(Correlation))
BURN_CORR$Significance <- ifelse(BURN_CORR$PValue <= 0.05, TRUE, FALSE)
BURN_CORR <- BURN_CORR[-1,]
rownames(BURN_CORR) <- NULL
kable(BURN_CORR, caption = "Table 2: Correlation Between H-2A Population and Fire/Heat Environmental Risk")

# Corrplot
COR_MTX <- round(cor(BURN, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "lightgrey",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 2,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  ) + 
  labs(title = "Figure 4: Correlation Between H-2A Population & Environmental Risk")
```


## Correlation between healthcare facilities per capita, migrant healthcare centers (not adjusted) and wildfire, drought, and heat wave exposure

These correlations are not very strong. Many of these correlations are due to the specific labeling of the healthcare centers, and because of this, there are less quantities of each type recorded. The figures below contain the correlation coefficients between healthcare and exposure.


```{r}
FIRE <- JOIN_CUT |> select(40:length(JOIN_CUT), matches("drought|wildfire|heat|heatwave"))
colnames(FIRE) <- gsub(" ", "", colnames(FIRE))
colnames(FIRE) <- gsub("(^|_)([a-z])", "\\1\\U\\2", colnames(FIRE), perl = T)
colnames(FIRE) <- gsub("\\$", "Dollars", colnames(FIRE))
colnames(FIRE) <- gsub("_", "", colnames(FIRE))


FIRE[] <- lapply(FIRE, as.numeric)
FIRE_HC <- FIRE |> select(3:20)
FIRE_HC <- cbind(FIRE_HC, Population2020 = JOIN$Population2020)
FIRE_HC <- FIRE_HC |>
  mutate(across(
    .cols = -c(MigrantHealthCenters, Population2020),
    .fns = ~ .x / Population2020
  ))
colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")] <-
  paste0(colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")], "PerCapita")
FIRE_HC <- FIRE_HC |> select(-Population2020)

FIRE_PRED <- FIRE |> select(21:length(FIRE))

FIRE_CORR <- data.frame(Response = character(),
                        Predictor = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (response in colnames(FIRE_HC)) {
  for( predictor in colnames(FIRE_PRED)) {
    FIRE_test <- cor.test(FIRE_HC[[response]], FIRE_PRED[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
    FIRE_CORR <- rbind(FIRE_CORR, data.frame(Response = response, Predictor = predictor, Correlation = round(FIRE_test$estimate, 3), PValue = round(FIRE_test$p.value, 3)))
  }
}

FIRE_CORR <- FIRE_CORR |> arrange(desc(Correlation))
FIRE_CORR$Significance <- ifelse(FIRE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(FIRE_CORR) <- NULL

kable(FIRE_CORR, caption = "Table 3: Correlation between healthcare facilities and Fire/Heat Environmental Risk")
```



## Correlation between farm output and wildfire, drought, and heat wave exposure


These correlations are also unsurprising. A firm's value is expected to be associated with their income/number of sales, so any location with high value will also have a high exposure score. The figures below contain the correlation coefficients between output and exposure.

```{r, fig.width = 8, fig.height = 10}
BLAZE_PRED <- JOIN_CUT |> select(matches("drought|wildfire|heat|heatwave"))
BLAZE_AGO <- JOIN_CUT |> select(intersect(colnames(AGO), colnames(JOIN_CUT))) |> select(-state, -county)
BLAZE_AGO[] <- lapply(BLAZE_AGO, as.numeric)
BLAZE_AGO[is.na(BLAZE_AGO)] <- 0
BLAZE <- cbind(BLAZE_PRED, BLAZE_AGO)

colnames(BLAZE) <- gsub("_", " ", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE), perl = TRUE)
colnames(BLAZE) <- gsub(" ", "", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\$", "Dollars", colnames(BLAZE))

colnames(BLAZE_AGO) <- gsub("_", " ", colnames(BLAZE_AGO))
colnames(BLAZE_AGO) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE_AGO), perl = TRUE)
colnames(BLAZE_AGO) <- gsub(" ", "", colnames(BLAZE_AGO))
colnames(BLAZE_AGO) <- gsub("\\$", "Dollars", colnames(BLAZE_AGO))

colnames(BLAZE_PRED) <- gsub("_", " ", colnames(BLAZE_PRED))
colnames(BLAZE_PRED) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE_PRED), perl = TRUE)
colnames(BLAZE_PRED) <- gsub(" ", "", colnames(BLAZE_PRED))
colnames(BLAZE_PRED) <- gsub("\\$", "Dollars", colnames(BLAZE_PRED))

BLAZE_AGO <- BLAZE_AGO |> select(-JOINCUTDollarsLatitude, -JOINCUTDollarsLongitude)

response <- character()
predictor <- character()

BLAZE_CORR <- data.frame(Predictor = character(),
                         Response = character(),
                         Correlation = numeric(),
                         PValue = numeric())

for (response in colnames(BLAZE_AGO)) {
  for (predictor in colnames(BLAZE_PRED)) {
    BLAZE_test <- cor.test(BLAZE[[response]], BLAZE[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
    BLAZE_CORR <- rbind(BLAZE_CORR, data.frame(Predictor = predictor, Response = response, Correlation = round(BLAZE_test$estimate, 3), PValue = round(BLAZE_test$p.value, 3)))
  }
}

BLAZE_CORR <- BLAZE_CORR |> arrange(desc(Correlation))
BLAZE_CORR$Significance <- ifelse(BLAZE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(BLAZE_CORR) <- NULL
BLAZE <- BLAZE |> select(-JOINCUTDollarsLatitude, -JOINCUTDollarsLongitude)
kable(BLAZE_CORR, caption = "Table 4: Correlation between Farm output and Wildfire/Drought/HeatWave Exposure")

# Corrplot
COR_MTX <- round(cor(BLAZE, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "lightgray",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 1.34,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  ) + 
  labs(title = "Figure 5: Correlation between Farm Output & Environmental Risk")
```

### Data Sources


1) H-2A Workers: [U.S. Department of Labor](https://www.uscis.gov/tools/reports-and-studies/h-2a-employer-data-hub)

2) Health Facilities: 

i) [Wyoming Department of Health](https://health.wyo.gov/aging/hls/healthcare-facility-directory/)

ii) [Utah Geospatial Resource Center](https://gis.utah.gov/data/health/health-care-facilities/)

iii) [South Dakota Department of Health](https://www.sdhls.org/verify/)

iv) [North Dakota Health and Human Services](https://www.hhs.nd.gov/)

v) [Montana Department of Health and Human Services](https://www.dphhs.mt.gov/qad/Licensure/)

vi) [Montana Primary Care Association](https://www.mtpca.org/wp-content/uploads/CHC_locations_MT_2022.pdf)

vii) [Colorado Department of Public Health and Environment](https://cdphe.colorado.gov/health-facilities/find-and-compare-facilities)

3) Migrant Health Centers: [National Center for Farmworker Health, Inc. (2023)](https://www.ncfh.org/migrant-health-centers.html)

4) Agricultural Census: [U.S. Department of Agriculture (2007-2023)](https://www.nass.usda.gov/Quick_Stats/CDQT/chapter/2/table/42/state/CO/year/2017/)

5) Maximum Temperature by County: [National Oceanic and Atmospheric Association (NOAA), 2023](https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/county/mapping/5/tmax/202208/1/value)

6) [National Risk Index](https://hicahs-centroid-collaboration-csurams.hub.arcgis.com/apps/1ff51926aada4a85a74ee89ba1a2539b/explore):  Federal Emergency Management Agency (ArcGIS Living Atlas)


### Appendix: 


```{r appendix, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(ggcorrplot)
library(leaflet)
library(leaflet.extras)
library(magrittr)
library(sf)
library(knitr)

opts_chunk$set(echo = F, eval = T, warning = F, message = F)

# Data:
H2A <- read_csv("~/internship/workspace/Written Datasets/h2a_by_county_new.csv") # Worker density
MHC <- read_csv("~/internship/workspace/migrant_health_centers_ncfh.csv") # Migrant health centers
ENV <- read_csv("~/internship/workspace/Written Datasets/disaster_cut_clean.csv") # Natural disaster and weather risk
AGO <- read_csv("~/internship/workspace/Written Datasets/ag_output_clean.csv") # Agricultural output
JUL23 <- read_csv("~/internship/HICAHS/Data/Heat_Ag_HumanRisk/CountyMaxTemp_JUL23.csv") # Maximum temperature data for July 2023
AUG23 <- read_csv("~/internship/HICAHS/Data/Heat_Ag_HumanRisk/CountyMaxTemp_AUG23.csv") # Maximum temperature data for August 2023

# Healthcare facilities in each state
# (Cannot bind due to differing number of columns)
COH <- read_excel("~/internship/workspace/Health Facility Data/ColoradoHealth24_xlsx.xlsx", sheet = 2)
WYH <- read_excel("~/internship/workspace/Health Facility Data/WyomingHealth24_xlsx.xlsx", sheet = 2)
SDH <- read_excel("~/internship/workspace/Health Facility Data/SouthDakotaHealth24_xlsx.xlsx", sheet = 2)
NDH <- read_excel("~/internship/workspace/Health Facility Data/NorthDakotaHealth24_xlsx.xlsx", sheet = 1)
UTH <- read_excel("~/internship/workspace/Health Facility Data/UtahHealth24_xlsx.xlsx", sheet = 2)
MTH <- read_excel("~/internship/workspace/Health Facility Data/MontanaHealth24_xlsx.xlsx", sheet = 1)

# Shape file for mapping
HICAHS_SHP <- invisible(st_read("~/internship/workspace/tl_2024_us_state/tl_2024_us_state.shp", quiet = TRUE))
COUNTY_SHP <- invisible(st_read("~/internship/workspace/tl_2024_us_county/tl_2024_us_county.shp", quiet = TRUE))

response <- character()
predictor <- character()

ENV$county[ENV$county == "lamoure"] <- "la moure"
ENV <- ENV |> select(-matches("coastal|hurricane|tsunami|volcanic", ignore.case = TRUE))
AGO[] <- lapply(AGO, function(col) gsub(",", "", col))

WYH$county <- gsub(" County", "", WYH$county)
MTH$county <- str_to_title(MTH$county)

COH$state <- str_to_title(COH$state)
WYH$state <- str_to_title(WYH$state)
SDH$state <- str_to_title(SDH$state)
NDH$state <- str_to_title(NDH$state)
UTH$state <- str_to_title(UTH$state)
MTH$state <- str_to_title(MTH$state)

JUL23$Name <- gsub(" County", "", JUL23$Name)
AUG23$Name <- gsub(" County", "", AUG23$Name)

JUL23 <- JUL23 |> rename(Jul2023_MaxTemp = Value)
AUG23 <- AUG23 |> rename(Aug2023_MaxTemp = Value)

JUL23 <- JUL23 |> mutate(year = 2023, month = "July", .before = Name)
AUG23 <- AUG23 |> mutate(year = 2023, month = "August", .before = Name)

JUL23 <- JUL23 |> rename(county = Name) |> rename(state = State) |> rename(`July1901-2000Mean` = `1901-2000 Mean`)
AUG23 <- AUG23 |> rename(county = Name) |> rename(state = State) |> rename(`Aug1901-2000Mean` = `1901-2000 Mean`)

# Column uniformity
H2A$county <- str_to_title(H2A$county)
MHC$county <- str_to_title(MHC$county)
ENV$county <- str_to_title(ENV$county)
AGO$county <- str_to_title(AGO$county)

H2A$state <- str_to_title(H2A$state)
MHC$state <- str_to_title(MHC$state)
ENV$state <- str_to_title(ENV$state)
AGO$state <- str_to_title(AGO$state)
AUG23$state <- str_to_title(AUG23$state)
JUL23$state <- str_to_title(JUL23$state)

# Adding latitude and longitude for mapping
MHC <- MHC |> separate(geolocation, into = c("latitude", "longitude"), sep = ",", convert = TRUE)

# Sum number of migrant health facilities per county
MHC_sum <- MHC |>
  group_by(state, county) |>
  summarise(MigrantHealthCenters = n(), .groups = "drop")


# Binding all data together into one data set to work with:
JOIN <- left_join(ENV, H2A, by = c("state", "county")) # Join H-2A worker totals and natural disaster data sets
JOIN <- left_join(JOIN, AGO, by = c("state", "county")) # Join agricultural output data set
JOIN <- left_join(JOIN, MHC_sum, by = c("state", "county")) # Join migrant health center counts
JOIN <- JOIN |>
  left_join(COH, by = c("state", "county")) |>
  left_join(WYH, by = c("state", "county")) |>
  left_join(SDH, by = c("state", "county")) |>
  left_join(NDH, by = c("state", "county")) |>
  left_join(UTH, by = c("state", "county")) |>
  left_join(MTH, by = c("state", "county"))

JOIN <- JOIN |> mutate(MigrantHealthCenters = ifelse(is.na(MigrantHealthCenters), 0, MigrantHealthCenters))
JOIN <- JOIN |> select(-grep("abbrev", colnames(JOIN), ignore.case = T))

JOIN <- JOIN |> mutate(Hospitals = coalesce(Hospitals.x, Hospitals.y, Hospitals.x.x, Hospitals.y.y, Hospitals.x.x.x, Hospitals.y.y.y))
JOIN <- JOIN |> select(-Hospitals.x, -Hospitals.y, -Hospitals.x.x, -Hospitals.y.y, -Hospitals.x.x.x, -Hospitals.y.y.y)

JOIN <- JOIN |> mutate(`RuralHealthClinics` = coalesce(`Rural_Clinics`, `Rural Health Clinics`, Rural_Health_Clinics.x, Rural_Health_Clinics.y))
JOIN <- JOIN |> select(-`Rural_Clinics`, -`Rural Health Clinics`, -Rural_Health_Clinics.x, -Rural_Health_Clinics.y)

JOIN <- JOIN |> mutate(`CriticalAccessHospitals` = coalesce(`Critical Access Hospitals`, Critical_Access_Hospitals))
JOIN <- JOIN |> select(-`Critical Access Hospitals`, -Critical_Access_Hospitals)

JOIN$county[JOIN$county == "lamoure"] <- "la moure"
#---------------------------------------------------------------------------------------------------------------#
JOIN_CUT <- JOIN |>
  select(
    "state",
    "county",
    "Population2020",
    "BuildingValue",
    "AgricultureValue",
    "Areasqmi",
    "H2A_workers",
    "JOIN_CUT$latitude",
    "JOIN_CUT$longitude",
    grep(
      "income|sales|farm_sales|income_net_|commodity_totals|crop_totals_sales|drought|wildfire|heat|heatwave",
      colnames(JOIN),
      ignore.case = TRUE
    ),
    349:length(JOIN)
  )

# Very simplified subset of risk variables
JOIN_CUT <- JOIN_CUT |> select(-grep("Expected|wheat|cattle|hogs|chickens|receipts|historic|index", colnames(JOIN_CUT), ignore.case = T))
# Filling NAs
JOIN_CUT[sapply(JOIN_CUT, is.numeric)] <- lapply(JOIN_CUT[sapply(JOIN_CUT, is.numeric)], function(x) ifelse(is.na(x), 0, x))
# Deleting empty columns
JOIN_CUT <- JOIN_CUT[, colSums(!is.na(JOIN_CUT)) > 0]

JOIN_CUT <- JOIN_CUT |>
  left_join(JUL23 |> select(state, county, Jul2023_MaxTemp, `July1901-2000Mean`), by = c("state", "county")) |>
  left_join(AUG23 |> select(state, county, Aug2023_MaxTemp, `Aug1901-2000Mean`), by = c("state", "county"))

JOIN_CUT <- JOIN_CUT |> arrange(state, county)


GGDAT1 <- MHC_sum |> 
  group_by(state) |> 
  summarise(total = sum(MigrantHealthCenters, na.rm = TRUE))

GGDAT1 <- rbind(GGDAT1, c("South Dakota", 0), c("Wyoming", 0))

GGPLOT1 <- ggplot(data = GGDAT1, aes(x = state, y = as.numeric(total))) + 
  geom_bar(stat = "identity", color = "black", fill = "darkgreen", width = .75) + 
  labs(x = "State", y = "Migrant Health Centers", title = "Figure 2: Migrant Health Centers Per State (NCFH 2023)") + 
  scale_y_continuous(breaks = seq(from = 0, to = 40, by = 2)) + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

GGDAT2 <- H2A |> 
  group_by(state) |> 
  summarise(total = sum(H2A_workers, na.rm = TRUE))

GGDAT1$total <- as.numeric(GGDAT1$total)
GGDAT2$total <- as.numeric(GGDAT2$total)

GGDAT1$state <- as.factor(GGDAT1$state)
GGDAT2$state <- as.factor(GGDAT2$state)

GGPLOT2 <- ggplot(data = GGDAT2, aes(x = state, y = as.numeric(total))) + 
  geom_bar(stat = "identity", color = "black", fill = "skyblue", width = .75) + 
  labs(x = "State", y = "Worker Count", title = "Figure 1: H-2A Workers") + 
  scale_y_continuous(breaks = seq(from = 0, to = max(GGDAT2$total), by = 500)) + 
  theme_minimal() + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

GGPLOT2
GGPLOT1



selected_states <- HICAHS_SHP[HICAHS_SHP$NAME %in% c("Wyoming", "Colorado", "Montana", "Utah", "North Dakota", "South Dakota"), ]
selected_counties <- COUNTY_SHP[COUNTY_SHP$STATEFP %in% selected_states$STATEFP, ] |> arrange(STATEFP, NAME)



JOIN_CUT_CLEAN <- JOIN_CUT |> filter(county != "Lamoure")
JOIN_CUT_sf <- st_as_sf(JOIN_CUT_CLEAN, coords = c("JOIN_CUT$longitude", "JOIN_CUT$latitude"), crs = 4326)


color_palette <- colorNumeric(palette = c("lightblue", "violet", "red"), domain = JOIN_CUT_sf$H2A_workers)
map <- leaflet(MHC) |>
  addTiles() |>
  setView(lng = -104.993498, lat = 42.468594, zoom = 5.45) |>
  addProviderTiles(providers$Esri.WorldTopoMap) |>
  addPolygons(data = selected_states,
              fill = FALSE,
              color = "black",
              weight = 0.5,
              opacity = 1) |>
  addPolygons(data = selected_counties,
              fillColor = ~color_palette(JOIN_CUT_sf$H2A_workers),
              fillOpacity = 0.7,
              color = "darkgrey",
              weight = 0.5,
              opacity = 1) |>
  addCircleMarkers(lng = ~longitude, lat = ~latitude,
                   fill = TRUE,
                   fillColor = "blue",
                   fillOpacity = 1,
                   stroke = T,
                   radius = 2) |>
  addLegend(pal = color_palette, values = JOIN_CUT_sf$H2A_workers, title = "H-2A Worker Population", position = "bottomright")

map



H2AMHC <- H2A |> left_join(MHC_sum, by = c("state", "county"))
H2AMHC$MigrantHealthCenters[is.na(H2AMHC$MigrantHealthCenters)] <- 0

MHC_CORR <- data.frame(Variable = character(),
                       Correlation = numeric(),
                       PValue = numeric() )

test <- cor.test(H2AMHC$H2A_workers, H2AMHC$MigrantHealthCenters, method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
MHC_CORR <- rbind(MHC_CORR, data.frame(Variable = "MigrantHealthCenters",
                                       Correlation = round(test$estimate, 3),
                                       PValue = round(test$p.value, 3 )))

MHC_CORR$Significance <- ifelse(MHC_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(MHC_CORR) <- NULL
kable(MHC_CORR, caption = "Table 1: Correlation between H-2A Population and Migrant Health Centers")



BURN <- JOIN_CUT |> select("H2A_workers", matches("drought|wildfire|heat|heatwave|Aug|Jul|July"))
BURN_CORR <- data.frame(Variable = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (predictor in colnames(BURN)) {
  test <- cor.test(BURN$H2A_workers, BURN[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
  BURN_CORR <- rbind(BURN_CORR, data.frame(Variable = predictor,
                                           Correlation = round(test$estimate, 3),
                                           PValue = round(test$p.value,3 )))
}

BURN_CORR <- BURN_CORR |> arrange(desc(Correlation))
BURN_CORR$Significance <- ifelse(BURN_CORR$PValue <= 0.05, TRUE, FALSE)
BURN_CORR <- BURN_CORR[-1,]
rownames(BURN_CORR) <- NULL
kable(BURN_CORR, caption = "Table 2: Correlation Between H-2A Population and Fire/Heat Environmental Risk")

# Corrplot
COR_MTX <- round(cor(BURN, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "lightgrey",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 2,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  ) + 
  labs(title = "Figure 4: Correlation Between H-2A Population & Environmental Risk")



FIRE <- JOIN_CUT |> select(40:length(JOIN_CUT), matches("drought|wildfire|heat|heatwave"))
colnames(FIRE) <- gsub(" ", "", colnames(FIRE))
colnames(FIRE) <- gsub("(^|_)([a-z])", "\\1\\U\\2", colnames(FIRE), perl = T)
colnames(FIRE) <- gsub("\\$", "Dollars", colnames(FIRE))
colnames(FIRE) <- gsub("_", "", colnames(FIRE))


FIRE[] <- lapply(FIRE, as.numeric)
FIRE_HC <- FIRE |> select(3:20)
FIRE_HC <- cbind(FIRE_HC, Population2020 = JOIN$Population2020)
FIRE_HC <- FIRE_HC |>
  mutate(across(
    .cols = -c(MigrantHealthCenters, Population2020),
    .fns = ~ .x / Population2020
  ))
colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")] <-
  paste0(colnames(FIRE_HC)[!colnames(FIRE_HC) %in% c("MigrantHealthWorkers", "Population2020")], "PerCapita")
FIRE_HC <- FIRE_HC |> select(-Population2020)

FIRE_PRED <- FIRE |> select(21:length(FIRE))

FIRE_CORR <- data.frame(Response = character(),
                        Predictor = character(),
                        Correlation = numeric(),
                        PValue = numeric() )

for (response in colnames(FIRE_HC)) {
  for( predictor in colnames(FIRE_PRED)) {
    FIRE_test <- cor.test(FIRE_HC[[response]], FIRE_PRED[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
    FIRE_CORR <- rbind(FIRE_CORR, data.frame(Response = response, Predictor = predictor, Correlation = round(FIRE_test$estimate, 3), PValue = round(FIRE_test$p.value, 3)))
  }
}

FIRE_CORR <- FIRE_CORR |> arrange(desc(Correlation))
FIRE_CORR$Significance <- ifelse(FIRE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(FIRE_CORR) <- NULL

kable(FIRE_CORR, caption = "Table 3: Correlation between healthcare facilities and Fire/Heat Environmental Risk")



BLAZE_PRED <- JOIN_CUT |> select(matches("drought|wildfire|heat|heatwave"))
BLAZE_AGO <- JOIN_CUT |> select(intersect(colnames(AGO), colnames(JOIN_CUT))) |> select(-state, -county)
BLAZE_AGO[] <- lapply(BLAZE_AGO, as.numeric)
BLAZE_AGO[is.na(BLAZE_AGO)] <- 0
BLAZE <- cbind(BLAZE_PRED, BLAZE_AGO)

colnames(BLAZE) <- gsub("_", " ", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE), perl = TRUE)
colnames(BLAZE) <- gsub(" ", "", colnames(BLAZE))
colnames(BLAZE) <- gsub("\\$", "Dollars", colnames(BLAZE))

colnames(BLAZE_AGO) <- gsub("_", " ", colnames(BLAZE_AGO))
colnames(BLAZE_AGO) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE_AGO), perl = TRUE)
colnames(BLAZE_AGO) <- gsub(" ", "", colnames(BLAZE_AGO))
colnames(BLAZE_AGO) <- gsub("\\$", "Dollars", colnames(BLAZE_AGO))

colnames(BLAZE_PRED) <- gsub("_", " ", colnames(BLAZE_PRED))
colnames(BLAZE_PRED) <- gsub("\\b([a-z])", "\\U\\1", colnames(BLAZE_PRED), perl = TRUE)
colnames(BLAZE_PRED) <- gsub(" ", "", colnames(BLAZE_PRED))
colnames(BLAZE_PRED) <- gsub("\\$", "Dollars", colnames(BLAZE_PRED))

BLAZE_AGO <- BLAZE_AGO |> select(-JOINCUTDollarsLatitude, -JOINCUTDollarsLongitude)

response <- character()
predictor <- character()

BLAZE_CORR <- data.frame(Predictor = character(),
                         Response = character(),
                         Correlation = numeric(),
                         PValue = numeric())

for (response in colnames(BLAZE_AGO)) {
  for (predictor in colnames(BLAZE_PRED)) {
    BLAZE_test <- cor.test(BLAZE[[response]], BLAZE[[predictor]], method = "spearman", use = "complete.obs", exact = F, conf.level = 0.95)
    BLAZE_CORR <- rbind(BLAZE_CORR, data.frame(Predictor = predictor, Response = response, Correlation = round(BLAZE_test$estimate, 3), PValue = round(BLAZE_test$p.value, 3)))
  }
}

BLAZE_CORR <- BLAZE_CORR |> arrange(desc(Correlation))
BLAZE_CORR$Significance <- ifelse(BLAZE_CORR$PValue <= 0.05, TRUE, FALSE)
rownames(BLAZE_CORR) <- NULL
BLAZE <- BLAZE |> select(-JOINCUTDollarsLatitude, -JOINCUTDollarsLongitude)
kable(BLAZE_CORR, caption = "Table 4: Correlation between Farm output and Wildfire/Drought/HeatWave Exposure")

# Corrplot
COR_MTX <- round(cor(BLAZE, method = "spearman", use = "complete.obs"), 2)
PMAT <- cor_pmat(COR_MTX)
ggcorrplot(
  COR_MTX,
  p.mat = PMAT,
  method = "square",
  type = "lower",
  outline.color = "lightgray",
  sig.level = 0.05,
  insig = "blank",
  lab = T,
  lab_size = 1.34,
  legend.title = "Correlation"
) +
  theme(
    axis.text.x = element_text(size = 6.5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6.5)
  ) + 
  labs(title = "Figure 5: Correlation between Farm Output & Environmental Risk")
```

